---
title: "Predicting Manner of Barbell Lifts from Accelerometers"
author: "chandu dugyala"
output: html_document
---

## Project Information

Personnel tracking devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here : http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

# Data:

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv



## Settingup environment for data analysis

Below R packages are used for this project,these packages must load before continuning on project 

```{r}
library(lattice)
library(ggplot2)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(36154)
```


### Loading data for analysis


```{r}
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(UrlTrain))
testing  <- read.csv(url(UrlTest))

inTrain  <- createDataPartition(training$classe, p=0.6, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)
dim(TestSet)
```
### Cleaning Data

Reducing the NZV, NA variables from data

```{r}
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
allNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, allNA==FALSE]
TestSet  <- TestSet[, allNA==FALSE]
```
# removing variables which doesn't make sense for prediction

```{r}
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
dim(TestSet)
```




## Machine learning approach

The goal of this project is to be able to predict the manner in which a barbell lift was done. I will use a random forest and Decision Tree methods make this prediction, and use best results to answer the quiz. 

### Random Forests Methos

```{r}
modelB1 <- randomForest(classe ~. , data=TrainSet)
```
# prediction on Test dataset

```{r}
predictionsB1 <- predict(modelB1, TestSet, type = "class")
```
```{r}
confusionMatrix(predictionsB1, TestSet$classe)
```

### Decision Tree Methos

```{r}
modelDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
fancyRpartPlot(modelDecTree)
```

# Prediction on Test dataset
```{r}
predictDecTree <- predict(modelDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree
```

### Generating files to submit assignmnet

```{r}
predictTEST <- predict(modelB1, newdata=testing)
predictTEST
```
